A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-01-31 sections/Metatheory.tex
transfer

\subsection{Transfer to the Declarative System}
\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}

The transfer rules are shown in Figure~\ref{fig:trans}.
The rules recursively replace the first existential variable to a well-scoped mono-type,
so the algorithmic worklist will finally become a declarative one.
In order to maintain well-formedness,
the substitution should affect all the judgments and term variable bindings in the worklist.

\begin{figure}[t]
\framebox{$\Gm \sto \Om$} $\Gm$ transfers to $\Om$. \jimmy{Always substitute from the very first existential variable. Types in contexts ($\Gm,x:A,\ldots$) should be taken care.}
\begin{gather*}
\inferrule*[right=$\mathtt{{\sto}}\Om$]
{~}
{\Om \sto \Om}
\quad
\inferrule*[right=$\mathtt{{\sto}\al}$]
{\Om\vdash\tau \\ \Om,[\tau/\al]\Gm \sto \Om}
{\Om,\al,\Gm \sto \Om}
\end{gather*}
\caption{Transfer Rules}
\label{fig:trans}
\end{figure}

\begin{figure}[t]
\framebox{$\Om \rto \Om'$} $\Om$ reduces to $\Om'$.
\begin{gather*}
\begin{aligned}
\Om,a &\rto \Om \qquad \Om,\al \rto \Om \qquad \Om,x:A \rto \Om\\
\Om\Vdash A\le B &\rto \Om \text{ when } \Om\vdash A\le B\\
\Om\Vdash e\Lto A &\rto \Om \text{ when } \Om\vdash e\Lto A\\
\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \Om\vdash e\To A\\
\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \Om\vdash \appInf{A}{e}{C}\\
\end{aligned}
\end{gather*}
\caption{Algorithmic Worklist Typing Rules}
\label{fig:alg:typing}
\end{figure}

\begin{lemma}[Insert]
If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
, then $\Gm_L, \al, \Gm_R \sto \Om$.
\end{lemma}
\begin{lemma}[Extract]
If $\Gm_L, \al, \Gm_R \sto \Om$
, then $\exists \tau$ s.t. $\Gm\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
\end{lemma}

>>>>

\subsection{Transfer to the Declarative System}
%\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}

\begin{figure}[t]
\framebox{$\Gm \sto \Om$} $\Gm$ transfers to $\Om$.
\begin{gather*}
\inferrule*[right=$\mathtt{{\sto}}\Om$]
{~}
{\Om \sto \Om}
\quad
\inferrule*[right=$\mathtt{{\sto}\al}$]
{\Om\vdash\tau \\ \Om,[\tau/\al]\Gm \sto \Om}
{\Om,\al,\Gm \sto \Om}
\end{gather*}
\caption{Transfer Rules}
\label{fig:trans}
\end{figure}

The transfer rules are shown in Figure~\ref{fig:trans}.
Rule $\mathtt{{\sto}\al}$ recursively replace the first existential variable to a well-scoped mono-type
($\Om$, as a declarative worklist, contains no existential variables),
so the algorithmic worklist will finally become a declarative one.
In order to maintain well-formedness,
the substitution should affect all the judgments and term variable bindings in the scope of $\al$.

The following lemmas~\ref{lem:insert} and \ref{lem:extract}
generalize Rule $\mathtt{{\sto}\al}$ from substituting the first existential variable
to substituting any existential variable.

\begin{lemma}[Insert]\label{lem:insert}
If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
, then $\Gm_L, \al, \Gm_R \sto \Om$.
\end{lemma}
\begin{lemma}[Extract]\label{lem:extract}
If $\Gm_L, \al, \Gm_R \sto \Om$
, then $\exists \tau$ s.t. $\Gm_L\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
\end{lemma}

\begin{figure}[t]
\framebox{$\Om \rto \Om'$} $\Om$ reduces to $\Om'$.
\begin{gather*}
\begin{aligned}
\Om,a &\rto \Om \qquad \Om \qquad \Om,x:A \rto \Om\\
\Om\Vdash A\le B &\rto \Om \text{ when } \Om\vdash A\le B\\
\Om\Vdash e\Lto A &\rto \Om \text{ when } \Om\vdash e\Lto A\\
\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \Om\vdash e\To A\\
\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \Om\vdash \appInf{A}{e}{C}\\
\end{aligned}
\end{gather*}
\caption{Declarative Worklist Typing}
\label{fig:decl:worklist}
\end{figure}

Figure~\ref{fig:decl:worklist} relates the worklist-style declarative system
to the original declarative system.
A successful reduction means a valid variable declaration,
a valid subtyping or checking judgment, or a successful guessing for inferences.
Additionally, judgment chains are unfolded to match the original declarative rules.
Although $\Om$ is defined differently from $\Psi$,
our declarative worklist $\Om$ contains a ordered collection of variable definitions,
which naturally fits declarative subtyping and typing.
